---
title: "single-cell-RNA-Seq-Data-Analysis-Methods"
author: "Mehadi Hasan"
date: "7/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "scRNA-Seq-Data-Analysis-Methods"
author: "Mehadi Hasan"
date: 11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**set the directory**
```{r}
getwd()
setwd()
```

**Load Libraries**
```{r}
library(SummarizedExperiment)
library(SingleCellExperiment)
```

**Load Datasets**
```{r}
targetspath <- "targetsPE.txt"                                                                                                                                                      
countpath <- "countDF.xls"                                                                                                                                              
targets <- read.delim(targetspath, comment.char = "#")                                                                                                                              
rownames(targets) <- targets$SampleName                                                                                                                                     
countDF <- read.delim(countpath, row.names=1, check.names=FALSE) 
```

**Create SummarizedExperiment**
```{r}
(se <- SummarizedExperiment(assays=list(counts=countDF), colData=targets))                                                                                                        
## coerce to SingleCellExperiment
(sce <- as(se, "SingleCellExperiment"))
```

**Create SingleCellExperiment directly**
```{r}
sce.direct <- SingleCellExperiment(assays=list(counts=countDF), colData=targets)
sce.direct
```

```{r}
# Prepare data for plotting with embedding methods

# The data are preprocessed (_e.g._normalized) to plot them with the run embedding functions from the scran and scater packages.

library(scran)
library(scater)

#lognormalized
sce <- logNormCounts(sce)
colLabels(sce) <- factor(colData(sce)$Factor) # This uses replicate info from above targets file as pseudo-clusters
```

**Embed with different methods and plot results**
```{r}
#(a) tSNE
sce <- runTSNE(sce)
reducedDimNames(sce)
plotTSNE(sce, colour_by="label", text_by="label")
```

**(b) MDS**
```{r}
sce <- runMDS(sce)
reducedDimNames(sce)
plotMDS(sce, colour_by="label", text_by="label")
```

**(c) UMAP**

```{r}
sce <- runUMAP(sce)
reducedDimNames(sce)
plotUMAP(sce, colour_by="label", text_by="label")
```

**(d) PCA**

```{r}
sce <- runPCA(sce) # gives a warning due to small size of data set but it still works 
reducedDimNames(sce)
plotPCA(sce, colour_by="label", text_by="label")
```

```{r}
# Multiple components can be plotted in a series of pairwise plots. When more than two components are plotted, the diagonal boxes in the scatter plot matrix show the density for each component.

sce <- runPCA(sce, ncomponents=20) # gives a warning due to small size of data set but it still works 
reducedDimNames(sce)
plotPCA(sce, colour_by="label", text_by="label", ncomponents = 4)
```

**scRNA-Seq data**

**Load scRNA-Seq data**

```{r}
#The scRNAseq package is used to load the scRNA-Seq data set from Xenopus tail directly into a SingleCellExperiment object (Aztekin et al. 2019).

library(scRNAseq)
sce <- AztekinTailData()
dim(sce)
colnames(sce)
rownames(sce)
```

**Prepare data for plotting with embedding methods**

```{r}
library(scran)
library(scater)

sce <- logNormCounts(sce)
clusters <- quickCluster(sce)
sce <- computeSumFactors(sce, clusters=clusters)
colLabels(sce) <- factor(clusters)
table(colLabels(sce))
```


```{r}
#To acclerate the testing performance of the following code, the size of the expression matrix is reduced to cell types with values => 10^4

filter <- colSums(assays(sce)$counts) >= 10^4
sce <- sce[, filter]
```


**Embed with different methods and plot results**

```{r}
# As under the bulk RNA-Seq section, the embedding results are sequentially appended to the SingleCellExperiment object, meaning one can use the plot function whenever necessary
```

**(a) tSNE**

```{r}
sce <- runTSNE(sce)
reducedDimNames(sce)
plotTSNE(sce, colour_by="label", text_by="label")
```

**(b) MDS**

```{r}
sce <- runMDS(sce)
reducedDimNames(sce)
plotMDS(sce, colour_by="label", text_by="label")
```

**(c) UMAP**

```{r}
sce <- runUMAP(sce)
reducedDimNames(sce)
plotUMAP(sce, colour_by="label", text_by="label")
```

**(d) PCA**

```{r}
#PCA result plotted for first two components

sce <- runPCA(sce) 
reducedDimNames(sce)
plotPCA(sce, colour_by="label", text_by="label")
```

```{r}
#Multiple components can be plotted in a series of pairwise plots. When more than two components are plotted, the diagonal boxes in the scatter plot matrix show the density for each component.

sce <- runPCA(sce, ncomponents=20) 
reducedDimNames(sce)
plotPCA(sce, colour_by="label", text_by="label", ncomponents = 4)
```

**save File as rds**

```{r}
saveRDS( sce, file= "C:/Users/saifmehadi/Documents/scRNA-Seq-Data-Analysis-Methods/output.rds")
```


